Work Fragmentation in Developer Interaction Data

### 2016-02-10 paper_2.tex
RQ5

In RQ3, we saw different patterns of activity around positive and negative interruptions. In particular, the edits per minute were higher after a positive interruption; on the contrary they were lower after a negative interruption. The effect is even more marked in the immediate minutes after the interruption, as can be seen in the Figure \ref{fig:box_recovery_effect_udc}.

@@

The presence of this pattern and the important difference observed in Figure \ref{fig:box_recovery_effect_udc} led us to investigate the effect more closely. To this aim, we performed an exploratory study of the activities performed by programmers during the recovery time of an interruption. This is possible due to the high variety of events present in the UDC dataset, which can tell us a lot more about the type of activities developers do. Indeed, the UDC records individual commands that the developers issued in the IDE. The commands are very descriptive in the kind of activity that they represent.

@@

To better understand the reasons of the difference between positive and negative interruptions, we did an analysis of the common events during recovery time, depending on the nature of the interruption. For this we created a new classification of events on a lower level that takes into consideration what kind of activity is performed by the programmer. The new classification contains the categories described in the Table \ref{tbl:new_events_udc}. This table also contains the frequency per minute of the more frequent event on that category.

@@

\begin{enumerate}
\item \emph{Negative} interruptions---having higher overall activity before the interruptions rather than after it, and a lower edit ratio as well---could be indicative of actual interruptions. That is, a programmer is interrupted in his normal course of work, and needs to rebuild it's mental context after the interruption, performing additional program comprehension before resuming programming.
\item On the other hand, \emph{Positive} interruptions---with higher activity after the interruption, and a higher edit ratio---could be indicative of information-seeking behavior. Programmers, noticing that they are encountering a problem that they are not able to fix on their own, may look for outside help (asking an expert, querying a question and answer site such as Stack Overflow, etc), and once a solution is found, resume their work in the IDE in a much more productive manner.
\item \RR{What about debugging sessions?}
\end{enumerate}

@@

After a positive interruptions the events belonging to the \emph{edit-text} category, the \emph{file} category, the \emph{refactoring} and the \emph{perspective} category, are more frequent than in the average. In addition, these categories of events are less frequent than the average after a negative interruptions. Of these four, the two categories with the largest differences between positive and negative interruptions are \emph{edit-text} and \emph{refactoring}. Both of these categories are much more clearly associated with edition activities than program comprehesion, and thus fit the previous hypothesis. Of the two other categories, \emph{file} management is also related to program edition activities, why the role of the \emph{perspective} category is still to clarify.

@@

After a positive interruption, we can see that the more common activities are text edition, refactoring, file management and, to a lesser extent, navigation around perspectives. On the other hand the navigation around classes, find and replace, searches, debug and clean and build are actions more common after a negative one. The navigation in the code and control version activities are more common after the recovery time, having a lower frequency during a positive o negative interruption. Of course, this exploratory study does not allow us to confirm these hypotheses. However, the evidence we have discovered so far does not disprove them, which means that our confidence in these hypotheses is higher as a consequence.

>>>>

In RQ3, we saw different patterns of activity around positive and negative interruptions. In particular, we observed that the edits per minute were higher after a positive interruption; on the contrary they were lower after a negative interruption. The effect is even more marked in the immediate minutes after the interruption, as can be seen in the Figure \ref{fig:box_recovery_effect_udc}.

@@

The presence of this pattern and the important difference observed in Figure \ref{fig:box_recovery_effect_udc} led us to investigate the effect more closely. To this aim, we performed an exploratory study of the activities performed by programmers during the recovery time of an interruption. This is possible due to the high variety of events present in the UDC dataset, which can tell us a lot more about the type of activities that developers do than the Mylyn dataset can. Indeed, the UDC records individual commands that the developers issued in the IDE. The commands are very descriptive in the kind of activity that they represent.

@@

%To better understand the reasons of the difference between positive and negative interruptions, we did an analysis of the common events during recovery time, depending on the nature of the interruption. For this we created a new classification of events on a lower level that takes into consideration what kind of activity is performed by the programmer. The new classification contains the categories described in the Table \ref{tbl:new_events_udc}. This table also contains the frequency per minute of the more frequent event on that category.

@@

\begin{enumerate}
\item \emph{Negative} interruptions---having higher overall activity before the interruptions rather than after it, and a lower edit ratio as well---could be indicative of actual interruptions. That is, a programmer is interrupted in his normal course of work, and needs to rebuild his or her mental context after the interruption, performing additional program comprehension before resuming programming.
\item On the other hand, \emph{Positive} interruptions---with higher activity after the interruption, and a higher edit ratio---could be indicative of information-seeking behavior. Programmers, noticing that they are encountering a problem that they are not able to fix on their own, may look for outside help (asking an expert, querying a question and answer site such as Stack Overflow, etc), and once a solution is found, resume their work in the IDE in a much more productive manner.
%\item \RR{What about debugging sessions?}
\end{enumerate}

@@

After positive interruptions the events belonging to the \emph{edit-text} category, the \emph{file} category, the \emph{refactoring} and the \emph{perspective} category, are more frequent than in the average. In addition, these categories of events are less frequent than the average after a negative interruptions. Of these four, the two categories with the largest differences between positive and negative interruptions are \emph{edit-text} and \emph{refactoring}. Both of these categories are much more clearly associated with edition activities than program comprehesion, and thus fit the previous hypothesis. Of the two other categories, \emph{file} management is also related to program edition activities, while the role of the \emph{perspective} category is still to clarify.

@@

After a positive interruption, we can see that the more common activities are text edition, refactoring, file management and, to a lesser extent, navigation around perspectives. On the other hand the navigation around classes, find and replace, searches, debug and clean and build are actions more common after a negative one. The navigation in the code and control version activities are more common after the recovery time, having a lower frequency during a positive o negative interruption.

Of course, this exploratory study does not allow us to confirm these hypotheses. However, the evidence we have discovered so far does not disprove them, and rather goes in the direction of our hypotheses, which means that our confidence in these hypotheses is higher as a consequence.
